---
title: "Lab 25: Chicago Crime Data II"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```


## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Chicago Crime Data

Today we are going to look at a fairly largerdataset. Each row
of the data refers to a single reported crime in the City of Chicago:

```{r}
crimes <- read_csv("https://statsmaths.github.io/stat_data/chi_crimes_2016.csv")
```

The available variable are:

- `area_number`: the community area code of the crime; a number from 1-77
- `arrest_flag`: whether the crime resulted in an arrest; 0 is false and 1 is true                
- `domestic_flag`: whether the crime is classified as a domestic offense; 0 is false and 1 is true                    
- `night_flag`: did the crime occur at night (9pm - 3am); 0 is false and 1 is true               
- `burglary`: was the crime classified as a burglary? 0 is false and 1 is true
- `theft`: was the crime classified as a theft? 0 is false and 1 is true
- `battery`: was the crime classified as a battery? 0 is false and 1 is true
- `damage`: was the crime classified as a damage? 0 is false and 1 is true
- `assault`: was the crime classified as an assault? 0 is false and 1 is true
- `deception`: was the crime classified as criminal deception? 0 is false and 1 is true
- `robbery`: was the crime classified as a robbery? 0 is false and 1 is true
- `narcotics`: was the crime classified as a narcotics violation? 0 is false and 1 is true

We also have metadata about each community area within Chicago as well.
We will see how to use these shortly.

```{r}
ca <- read_csv("https://statsmaths.github.io/stat_data/chicago_meta.csv")
```

- `area_number`: the community area code; a number from 1 to 77
- `area_name`: popular name of the community area
- `median_age`: the median age of all residents in the community area
- `num_households`: total number of households
- `family_households`: percentage of households classified as a `family'
                     (domestic partners, married couples, and one or more
                     parents with children)
- `family_w_kids`: percentage of households with children under the age of 18                     
- `owner_ratio`: ratio of households that own or mortgage their primary residence                    
- `mean_travel_time`: average commute time
- `percent_walk`: percentage of commuters who walk to work (0-100)
- `median_income`: median household income
- `perc_20_units`: percentage of residential buildings with 20 or more units


As before, it will be useful to group summarize the dataset and join the
metadata. Do that here to construct the dataset `crimes_ca`:

```{r}
crimes_ca <- group_summarize(crimes, area_number)
crimes_ca <- left_join(crimes_ca, ca)
```

### Example

Last time, I said that it may be useful to summarize the dataset it in
a different way. This will be a powerful tool in Project D. To test your
knowledge of what this means, look at this dataset and plot:

```{r}
crimes_time_ca <- group_summarize(crimes, area_number, night_flag)
crimes_time_ca <- left_join(crimes_time_ca, ca, by = "area_number")
crimes_time_ca$burglary_rate <- crimes_time_ca$burglary_sum / crimes_time_ca$num_households * 1000

ggplot(crimes_time_ca, aes(median_income, burglary_rate)) +
  geom_point(aes(color = factor(night_flag))) +
  geom_smooth(aes(color = factor(night_flag)), method = "lm") +
  scale_color_viridis(discrete = TRUE)
```

How many observations are there in the data? What does one row represent?
How do you interpret the plot? Do the two slopes differ; what does this
mean in practical terms? 

There are 154 observations. Each row represents a particular area either at night or during the day. The higher the income, the less total burglaries, and also the less burglaries that take place at night. The slopes mean that income has less of an impact on the burglaries taking place during the day. 

### Prompts

Address one of the following prompts. Your end goal should be to produce
a data visualization (make sure to include labels) and a 1-2 sentence thesis
statement. Straight from Wikipedia, a thesis statement: "offers a concise
summary of the main point or claim of the essay, research paper, ect."
Note, these questions are jumping off points. You can deviate or extend
from them if something related strikes your interest.


2. What proportion of narcotics violations result in an arrest by neighborhood?
Do any features of a neighborhood seem to effect this rate? If you see any
outliers try to document and explain these.


### Exploratory Analysis

Write you exploratory analysis here:

```{r}
narcotics_arrest <- group_summarize(crimes, narcotics, area_number)
narcotics_arrest <- left_join(narcotics_arrest, ca, by = "area_number")

narcotics_crimes <- filter (narcotics_arrest, narcotics == 1)

ggplot (narcotics_crimes, aes (arrest_flag_mean, area_number)) +
  geom_point (aes(factor = narcotics)) +
  geom_smooth(aes(color = factor(narcotics)), method = "lm") +
  scale_color_viridis(discrete = TRUE) +
  geom_text (aes (label = area_name))
```

The vast majority of narcotics violations result in arrests 100% of the time, though there are a few outliers, most noteably the neighborhoods of Archer Heights, Irving Park, and Near West Side, where the arrest for narcotics is between 95.5% and 98.5%. This difference in arrest percentage is due to the changing affluence of the inhabitants of the neighborhoods, where less affluent neighborhoods will have more arrests. 
